\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-04-25 forwards-and-backwards.tex
removed general kernel from PME

\comment{TODO: ALISON SAYS
The proof of Lemma 6.7 is now problematic. We'll be hard pushed to find a sequence of functions increasing to x, or indeed -x. Easily fixed of course, but I think that we should anyway change the condition of the theorem to be a bit more general, so say that we assume that $\langle f(x),\eta^N_0(dx)\rangle$  is uniformly bounded for some non-negative, twice differentiable with uniformly bounded first and second derivatives, function $f$ with $f(x)\to\infty$ as $x\to\infty$ and then use Lemma 6.1 and Gronwall to deduce $\langle f,\eta_t\rangle$ has bounded mean for any limit point and so Markov's inequality tells us no mass escaped to infinity.
}

\comment{``reaction diffusion equation'' or ``reaction-diffusion equation''?}

@@

We introduce a broad class of mechanistic spatial models to describe how spatially 
heterogeneous populations live, die, and reproduce.
Individuals are represented by points of a point measure,
whose birth and death rates can depend 
both on spatial position and local population density, defined at a location
to be the convolution of the point measure
with a suitable non-negative integrable kernel
centred on that location. We pass to three different scaling limits:
an interacting superprocess, a nonlocal partial differential equation (PDE), 
and a classical PDE. 
The classical PDE is obtained both by a two-step convergence argument, in which we 
first scale time and population size and
pass to the nonlocal PDE, 
and then scale the kernel that 
determines local population density; and 
in the important special case in which the limit is a 
reaction diffusion equation, directly 
by simultaneously scaling the kernel width,
timescale and population size in our individual based model.

@@

It is of interest to understand under what conditions we can replace the two-step limiting
process described above by one in which we simultaneously scale the 
kernels
and the other parameters in 
our population model to arrive at the PDE limit.
This is mathematically much more challenging, but we establish such 
one-step convergence in cases for which the limit is a classical 
reaction diffusion equation of the form~(\ref{semilinear PDE}) with 
$\DG=\Delta$,
and $\rho_F$ is a Gaussian density. We allow a wide class of reaction terms, $F$, so
that the Fisher-KPP equation 
(that is equation~(\ref{semilinear PDE}) with $\DG=\Delta$ and $F(\varphi)=1-\varphi$)
emerges as a special case.

Such results on (one-step) convergence to reaction diffusion 
equation limits have been
achieved for a variety of interacting particle systems. 
Following the now classical contributions 
of~\citet{demasi/ferrari/lebowitz:1986, demasi/presutti:1991, oelschlaeger:1985},
much of this work has focused on lattice based models with one particle
per site, or on systems with a fixed number, $N$,
of interacting diffusions as $N\to\infty$.
For systems of proliferating particles,
as considered for example by 
\citet{oelschlaeger:1989, flandoli/leimbach/olivera:2019, flandoli/huang:2021}, 
an additional challenge (also apparent in our models), is 
the control of concentration of particles. 
We follow 
\citet{oelschlaeger:1989, flandoli/leimbach/olivera:2019} in considering
`moderate interactions', meaning that
the number of individuals in the neighbourhood over which we measure local
population density tends to infinity, whereas 
\citet{flandoli/huang:2021} also consider the situation in which
that number remains finite. 
We refer to \citet{flandoli/huang:2021} for a more thorough 
literature review, 
but note that both our model and scaling differ from those considered
in the body of work discussed there: whereas in those settings, the only 
scalings are the number of particles in the system and the size of the
neighbourhood over which individuals interact with one another, in
keeping with the vast literature on continuous state branching models, we 
also scale time and so must ensure that births are 
adequately compensated by deaths to prevent the population from exploding.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
    Suppose that Assumptions~\ref{def:model_setup} hold and, 
    further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
    purely atomic measures, with $\eta_0^N$ comprised of atoms of mass $1/N$,
    which is tight in $\measures$ and such that
    $\langle |x|, \eta_0^N(dx)\rangle<C<\infty$
    with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    If $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\paragraph{Two-step convergence to PDE:}
Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
we can choose our kernels $\rho_\gamma$, $\rho_r$, and $\rho_F$
in such a way that they depend only on the population in a region close to $x$,
and so we expect that under rather general conditions
solutions of the nonlocal PDE will be close to the corresponding 
classical PDE.
The following propositions provide two concrete situations in which this is true.
In the first, the PDE is a reaction diffusion equation, and we borrow
an idea from~\cite{penington:2017} to express the solutions to both the nonlocal
equation and the classical PDE through a Feynman-Kac formula.

@@

\comment{TODO: restrict to Gaussian case}
\begin{proposition}
	\label{nonlocalPME to PME}
Suppose that
we can write $\rho_\gamma=\zeta*\check{\zeta}$, where $\check{\zeta}(x)=\zeta(-x)$ and
$\zeta\in\mathcal{S}(\IR^d)$ (the Schwartz space of rapidly decreasing functions).
Furthermore, suppose that %$\psi_0^\epsilon\in L^1(\IR^d)$, 
$\psi_0^\epsilon\geq 0$ is such that
there exist $\lambda\in (0,1)$ and $C\in (0,\infty)$ (independent
of $\epsilon$) such that 
\begin{equation*}
\int\exp(\lambda \|x\|)\psi_0^\epsilon(x)dx<C,\quad\mbox{ and }
\sup_\epsilon\int \psi_0^\epsilon|\log \psi_0^\epsilon|dx<\infty,
\end{equation*}
with
$\psi_0^\epsilon\rightharpoonup \psi_0$ as $\epsilon\to 0$. Then
writing $\psi^\epsilon$ for the solution to~(\ref{mollified equation})
on $[0,T]\times \IR^d$ with
initial condition $\psi_0^\epsilon$,
$\psi^\epsilon\rightharpoonup \psi$ as $\epsilon\to 0$ where
$\psi\in L_t^2(H^1)\cap C_t(L^1)$, $\int \psi|\log \psi| dx<\infty$, and
$\psi$ solves~(\ref{PME}) on $[0,T]\times \IR^d$. 
\end{proposition}
The example that we have in mind for the kernel $\rho_\gamma$ is a Gaussian
kernel.
\begin{remark}
Although it seems hard to formulate an all-encompassing result,
Propositions~\ref{prop:nonlocal_to_local} and~\ref{nonlocalPME to PME} are by no
means exhaustive. When the scaling limit is deterministic, one can 
expect analogous results under rather general conditions. However, when the limit
points are stochastic, they resemble ``nonlinear superprocesses'' and so one cannot
expect a density with respect to Lebesgue measure in $d\geq 2$. It is then not
reasonable to 
expect to be able to make sense of the limit if we scale the kernels in this way.
Moreover, in one dimension, where the classical superprocess does have a density with 
respect to Lebesgue measure, the form of~(\ref{eqn:limiting_mgale_problem})
suggests that even if one can remove the local averaging from $\gamma$, it
will be necessary
to retain averaging of $r$ in order to obtain a well-defined limit. 
\end{remark}

@@

Equation~(\ref{eqn:nonlocal_pde}) is a nonlocal version of a reaction diffusion equation;
the diffusion is nonlinear if $\gamma$ depends on population density:
in other words, if the diffusivity of the population depends on the population density.
Passing to the classical limit, we recover equations like~(\ref{PME}).
Such equations are widely used in a number of contexts in biology in which
motility within a population varies with population density.
For example, density dependent dispersal is a common feature in spatial
models in ecology, eukaryotic cell biology, and avascular tumour growth;
see~\cite{sherratt:2010} and references therein for further discussion. 
In particular, it has been suggested as a model for
the expansion of a certain type of bacteria % of the type {\em Paenbacillus dendritiformis}
on a thin layer of agar in a petri dish 
\citep{cohen/golding/kozlovsky/benjacob/ron:1999}. 
We shall pay particular attention to the case in which the equation can be 
thought of as modelling the density of an expanding population. 
We focus instead on the monostable reaction of~(\ref{PME}).

@@

In this section we present formal proofs of the first two of our
three scaling limits.
%following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain (both
stochastic and deterministic) limits in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
%In subsequent sections (Sections~\ref{subsec:one step convergence proof} and~\ref{sec:lookdown_proofs})
%we prove our third scaling limit, in which 
%we pass from our population model 
%directly to a reaction diffusion equation without going via a nonlocal limit,
%and finally our results concerning both convergence of the lookdown representation
%and characterization of the lineage process in the deterministic case.
\comment{FROM ALISON: slightly reworded - deleted bit about onestep convergence
on the basis that we already point to that in the introduction}

@@

\begin{proof}[Sketch]
To avoid the idea being buried in notation, we sketch the proof in the 
special case of $d=1$.
We take a sequence of test functions $f_n$ increasing to the function $f(x)=x$
and having uniformly bounded first and second derivatives,
so that there is a (single) $C$
from Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
such that $\gamma(x,\eta) B_{f_n}^\theta(x,\eta) \le C(1 + f_n(x))$
for all $x$, $\eta$, and $f_n$.
Then, just as we arrived at equation~\eqref{bound on intfdeta},
\begin{align*}
    \IE[\langle f_n(x), \eta^N_t(dx) \rangle]
    &\le
    \IE[\langle f_n(x), \eta^N_0(dx) \rangle]
    +
    C \int_0^t \IE\Big[\Big\langle f_n(x), \eta^N_s(dx) \Big] \Big\rangle ds ,
\end{align*}
with the same constant for all $n$ and all $N$.
Gronwall's inequality then implies that 
$\IE[\langle f_n, \eta^N_t\rangle] \le C'$
for some $C'$ independent of $n$, $N$, and $t \in [0,T]$.
By first taking $N \to \infty$ and then $n \to \infty$,
we find that $\IE[\langle x, \eta_t(dx)\rangle] \le C'$
for $t\in [0,T]$, and an application of Markov's inequality
tells us that for any $\delta > 0$,
$\IP[\eta_t([R,\infty])>\delta]\to 0$ as $R\to\infty$.
The same argument with $f_n(x)$ increasing to $-x$ completes the proof.
\end{proof}

@@

It is natural to conjecture that when the limit of the rescaled 
population process that we obtained in the previous section solves a nonlocal
PDE, if we further scale the kernels 
$\rho_r$, $\rho_\gamma$, and $\rho_F$ by setting 
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon$, as $\epsilon\to 0$, the 
corresponding solutions should converge to a limiting
population density that solves the corresponding ``classical'' PDE. We verify
this in two examples; in the first the nonlocal equation is a reaction diffusion equation
with the ``nonlocality'' only appearing in the 
reaction term; in the second the nonlocal PDE is a special 
case of a nonlinear porous medium equation. These, in particular,
capture the examples that we explored in 
Section~\ref{ancestral lineages for travelling waves}.

@@

In this subsection we prove Proposition~\ref{nonlocalPME to PME}.
To ease notation, we drop the subscript $\gamma$ on $\rho_\gamma$ and we present 
the proof in $d=1$ (although we retain the notation $\nabla$).
It should be clear that it extends almost without change
to higher dimensions.

Recall that 
we are concerned with non-negative solutions to the equation~(\ref{mollified equation}):
\begin{equation*}
	\partial_t\psi_t^\epsilon(x) =
	\Delta\left(\psi_t^\epsilon \, \rho^\epsilon *\psi_t^\epsilon\right)(x)
	+\psi_t^\epsilon(x)\left(1-\rho^\epsilon *\psi_t^\epsilon(x) 
\right),
\end{equation*}
and 
we assume that $\rho=\zeta*\check{\zeta}$ with $\zeta$ a rapidly decreasing function. 
The example we have in mind is $\zeta$ (and therefore $\rho$) being the density of
a mean zero Gaussian random variable.  
We shall prove that under the 
assumptions of Proposition~\ref{nonlocalPME to PME},
as $\epsilon\to 0$, we have
convergence to the solution to the porous medium equation with
logistic growth,
equation~\eqref{PME1}:
\begin{equation*}
	\partial_t \psi_t(x)=
	\Delta\left(\psi_t^2\right)(x)
	+\psi_t(x)\left(1-\psi_t(x)\right).
\end{equation*}

We work on the time interval $[0,T]$.
We will require a lower bound on $\int \psi_t^\epsilon(x)\log \psi_t^\epsilon(x) dx$ 
which we record as a lemma.
\begin{lemma}
\label{lower bound on ulogu}
	Suppose that there exists $\lambda\in (0,1)$ and $C<\infty$, both independent of
$\epsilon$, such that $\int \exp(\lambda |x|) \psi_0^\epsilon(x) dx<C$.
Then there exists
a constant $K<\infty$, independent of $\epsilon$, such that  
 $\int \psi_t^\epsilon(x)\log \psi_t^\epsilon (x)dx>-K$ for all $t\in [0,T]$. 
\end{lemma}
\begin{proof}
First observe that, since $x\log x$ is bounded below, 
$\int_{-1}^1\psi_t^\epsilon(x)\log\psi_t^\epsilon(x) dx$ is bounded below,
and recall that $\psi_t^\epsilon(x) \ge 0$.

@@

\begin{proof}[Proof of Proposition~\ref{nonlocalPME to PME}]
First observe that
\begin{multline*}
	\int \psi_t^\epsilon(x)\, \rho^{\epsilon}*\psi_t^\epsilon(x) dx 
=
\int\int\int \psi_t^\epsilon(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z) dz dy dx 
\\
=
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{z})
\psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z}) d\tilde{z} d\tilde{y} d \tilde{x}
	=\int\left(\zeta^{\epsilon}*\psi_t^\epsilon(x)\right)^2 dx,
\end{multline*}
where we have set $\tilde{x}=x-z$, $\tilde{y}=y-z$, $\tilde{z}=-z$.

@@

In this section we prove Theorem~\ref{thm:local_convergence},
which simultaneously scales the parameters $\theta$ and $N$ and the interaction 
kernel $\rho_F$ in the special case in which $r\equiv 1\equiv\gamma$, $q_{\theta}(x,dy)$
is isotropic with zero mean, the kernel
$\rho_F$ is Gaussian, and the scaling 
limit is a reaction diffusion equation.

>>>>

@@

We introduce a broad class of mechanistic spatial models to describe how spatially 
heterogeneous populations live, die, and reproduce.
Individuals are represented by points of a point measure,
whose birth and death rates can depend 
both on spatial position and local population density, defined at a location
to be the convolution of the point measure
with a suitable non-negative integrable kernel
centred on that location. We pass to three different scaling limits:
an interacting superprocess, a nonlocal partial differential equation (PDE), 
and a classical PDE. 
The classical PDE is obtained both by a two-step convergence argument, in which we 
first scale time and population size and
pass to the nonlocal PDE, 
and then scale the kernel that 
determines local population density; and 
in the important special case in which the limit is a 
reaction-diffusion equation, directly 
by simultaneously scaling the kernel width,
timescale and population size in our individual based model.

@@

It is of interest to understand under what conditions we can replace the two-step limiting
process described above by one in which we simultaneously scale the 
kernels
and the other parameters in 
our population model to arrive at the PDE limit.
This is mathematically much more challenging, but we establish such 
one-step convergence in cases for which the limit is a classical 
reaction-diffusion equation of the form~(\ref{semilinear PDE}) with 
$\DG=\Delta$,
and $\rho_F$ is a Gaussian density. We allow a wide class of reaction terms, $F$, so
that the Fisher-KPP equation 
(that is equation~(\ref{semilinear PDE}) with $\DG=\Delta$ and $F(\varphi)=1-\varphi$)
emerges as a special case.

Such results on (one-step) convergence to reaction-diffusion 
equation limits have been
achieved for a variety of interacting particle systems. 
Following the now classical contributions 
of~\citet{demasi/ferrari/lebowitz:1986, demasi/presutti:1991, oelschlaeger:1985},
much of this work has focused on lattice based models with one particle
per site, or on systems with a fixed number, $N$,
of interacting diffusions as $N\to\infty$.
For systems of proliferating particles,
as considered for example by 
\citet{oelschlaeger:1989, flandoli/leimbach/olivera:2019, flandoli/huang:2021}, 
an additional challenge (also apparent in our models), is 
the control of concentration of particles. 
We follow 
\citet{oelschlaeger:1989, flandoli/leimbach/olivera:2019} in considering
`moderate interactions', meaning that
the number of individuals in the neighbourhood over which we measure local
population density tends to infinity, whereas 
\citet{flandoli/huang:2021} also consider the situation in which
that number remains finite. 
We refer to \citet{flandoli/huang:2021} for a more thorough 
literature review, 
but note that both our model and scaling differ from those considered
in the body of work discussed there: whereas in those settings, the only 
scalings are the number of particles in the system and the size of the
neighbourhood over which individuals interact with one another, in
keeping with the vast literature on continuous state branching models, we 
also scale time and so must ensure that births are 
adequately compensated by deaths to prevent the population from exploding.

@@

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
    Suppose that Assumptions~\ref{def:model_setup} hold and, 
    further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
    purely atomic measures, with $\eta_0^N$ comprised of atoms of mass $1/N$,
    which is tight in $\measures$.
    Also assume there exists a nonnegative $f_0 \in C(\IR^d)$
    with uniformly bounded first and second derivatives
    (i.e., with $\sup_x \sup_{\|z\|=1} \sum_i \partial_{x_i} f_0(x) z_i$
    and $\sup_x \sup_{\|z\|=1} \sum_{ij} \partial_{x_i x_j} f_0(x) z_i z_j$ both finite)
    and $f_0(x) \to \infty$ as $|x| \to \infty$ for which
    $\langle f_0(x), \eta_0^N(dx)\rangle<C<\infty$
    for some $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    If $\alpha = 0$ the limit is deterministic.
\end{theorem}

@@

\paragraph{Two-step convergence to PDE:}
Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
we can choose our kernels $\rho_\gamma$, $\rho_r$, and $\rho_F$
in such a way that they depend only on the population in a region close to $x$,
and so we expect that under rather general conditions
solutions of the nonlocal PDE will be close to the corresponding 
classical PDE.
The following propositions provide two concrete situations in which this is true.
In the first, the PDE is a reaction-diffusion equation, and we borrow
an idea from~\cite{penington:2017} to express the solutions to both the nonlocal
equation and the classical PDE through a Feynman-Kac formula.

@@

\begin{proposition}
	\label{nonlocalPME to PME}
Suppose that
% we can write $\rho_\gamma=\zeta*\check{\zeta}$, where $\check{\zeta}(x)=\zeta(-x)$ and
% $\zeta\in\mathcal{S}(\IR^d)$ (the Schwartz space of rapidly decreasing functions).
$\rho_\gamma = p_{\epsilon^2_\gamma}$,
and that %$\psi_0^\epsilon\in L^1(\IR^d)$, 
$\psi_0^\epsilon\geq 0$ is such that
there exist $\lambda\in (0,1)$ and $C\in (0,\infty)$ (independent
of $\epsilon$) such that 
\begin{equation*}
\int\exp(\lambda \|x\|)\psi_0^\epsilon(x)dx<C,\quad\mbox{ and }
\sup_\epsilon\int \psi_0^\epsilon|\log \psi_0^\epsilon|dx<\infty,
\end{equation*}
with
$\psi_0^\epsilon\rightharpoonup \psi_0$ as $\epsilon\to 0$. Then
writing $\psi^\epsilon$ for the solution to~(\ref{mollified equation})
on $[0,T]\times \IR^d$ with
initial condition $\psi_0^\epsilon$,
$\psi^\epsilon\rightharpoonup \psi$ as $\epsilon\to 0$ where
$\psi\in L_t^2(H^1)\cap C_t(L^1)$, $\int \psi|\log \psi| dx<\infty$, and
$\psi$ solves~(\ref{PME}) on $[0,T]\times \IR^d$. 
\end{proposition}
The example that we have in mind for the kernel $\rho_\gamma$ is a Gaussian
kernel.
\begin{remark}
Although it seems hard to formulate an all-encompassing result,
Propositions~\ref{prop:nonlocal_to_local} and~\ref{nonlocalPME to PME} are by no
means exhaustive. When the scaling limit is deterministic, one can 
expect analogous results under rather general conditions. However, when the limit
points are stochastic, they resemble ``nonlinear superprocesses'' and so one cannot
expect a density with respect to Lebesgue measure in $d\geq 2$. It is then not
reasonable to 
expect to be able to make sense of the limit if we scale the kernels in this way.
Moreover, in one dimension, where the classical superprocess does have a density with 
respect to Lebesgue measure, the form of~(\ref{eqn:limiting_mgale_problem})
suggests that even if one can remove the local averaging from $\gamma$, it
will be necessary
to retain averaging of $r$ in order to obtain a well-defined limit. 
\end{remark}

@@

Equation~(\ref{eqn:nonlocal_pde}) is a nonlocal version of a reaction-diffusion equation;
the diffusion is nonlinear if $\gamma$ depends on population density:
in other words, if the diffusivity of the population depends on the population density.
Passing to the classical limit, we recover equations like~(\ref{PME}).
Such equations are widely used in a number of contexts in biology in which
motility within a population varies with population density.
For example, density dependent dispersal is a common feature in spatial
models in ecology, eukaryotic cell biology, and avascular tumour growth;
see~\cite{sherratt:2010} and references therein for further discussion. 
In particular, it has been suggested as a model for
the expansion of a certain type of bacteria % of the type {\em Paenbacillus dendritiformis}
on a thin layer of agar in a petri dish 
\citep{cohen/golding/kozlovsky/benjacob/ron:1999}. 
We shall pay particular attention to the case in which the equation can be 
thought of as modelling the density of an expanding population. 
We focus instead on the monostable reaction of~(\ref{PME}).

@@

In this section we present formal proofs of the first two of our
three scaling limits.
%following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain (both
stochastic and deterministic) limits in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction-diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
%In subsequent sections (Sections~\ref{subsec:one step convergence proof} and~\ref{sec:lookdown_proofs})
%we prove our third scaling limit, in which 
%we pass from our population model 
%directly to a reaction-diffusion equation without going via a nonlocal limit,
%and finally our results concerning both convergence of the lookdown representation
%and characterization of the lineage process in the deterministic case.
\comment{FROM ALISON: slightly reworded - deleted bit about onestep convergence
on the basis that we already point to that in the introduction}

@@

\begin{proof}[Sketch]
Take $f_0(x)$ as in the statement of Theorem~\ref{thm:nonlocal_convergence},
i.e., $f_0$ is nonnegative, grows to infinity as $x \to \infty$,
has uniformly bounded first and second derivatives,
and has $\langle f_0, \eta^N_0\rangle$ uniformly bounded in $N$.
We take a sequence of test functions $f_n$ that increase to the function $f_0$
and having uniformly bounded first and second derivatives,
so that there is a (single) $C$
from Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
such that $\gamma(x,\eta) B_{f_n}^\theta(x,\eta) \le C(1 + f_n(x))$
for all $x$, $\eta$, and $f_n$.
Then, just as we arrived at equation~\eqref{bound on intfdeta},
\begin{align*}
    \IE[\langle f_n(x), \eta^N_t(dx) \rangle]
    &\le
    \IE[\langle f_n(x), \eta^N_0(dx) \rangle]
    +
    C \int_0^t \IE\Big[\Big\langle f_n(x), \eta^N_s(dx) \Big] \Big\rangle ds ,
\end{align*}
with the same constant for all $n$ and all $N$.
Gronwall's inequality then implies that 
$\IE[\langle f_n, \eta^N_t\rangle] \le C'$
for some $C'$ independent of $n$, $N$, and $t \in [0,T]$.
By first taking $N \to \infty$ and then $n \to \infty$,
we find that $\IE[\langle f_0, \eta_t(dx)\rangle] \le C'$
for $t\in [0,T]$, and since $f_0 \to \infty$ as $|x| \to \infty$,
an application of Markov's inequality
tells us that for any $\delta > 0$,
$\IP[\eta_t(\{x : \|x\| > R\})>\delta]\to 0$ as $R\to\infty$.
\end{proof}

@@

It is natural to conjecture that when the limit of the rescaled 
population process that we obtained in the previous section solves a nonlocal
PDE, if we further scale the kernels 
$\rho_r$, $\rho_\gamma$, and $\rho_F$ by setting 
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon$, as $\epsilon\to 0$, the 
corresponding solutions should converge to a limiting
population density that solves the corresponding ``classical'' PDE. We verify
this in two examples; in the first the nonlocal equation is a reaction-diffusion equation
with the ``nonlocality'' only appearing in the 
reaction term; in the second the nonlocal PDE is a special 
case of a nonlinear porous medium equation. These, in particular,
capture the examples that we explored in 
Section~\ref{ancestral lineages for travelling waves}.

@@

In this subsection we prove Proposition~\ref{nonlocalPME to PME}.
To ease notation, we present
the proof in $d=1$ (although we retain the notation $\nabla$).
However to recall the dependence on $\epsilon$ we write $\rho^\epsilon$ for $\rho_\gamma$
(which recall for this Proposition we suppose is Gaussian
and equal to $p_{\epsilon^2_\gamma}$).
It should be clear that it extends almost without change
to higher dimensions.

Recall that 
we are concerned with non-negative solutions to the equation~(\ref{mollified equation}):
\begin{equation*}
	\partial_t\psi_t^\epsilon(x) =
	\Delta\left(\psi_t^\epsilon \, \rho^\epsilon *\psi_t^\epsilon\right)(x)
	+\psi_t^\epsilon(x)\left(1-\rho^\epsilon *\psi_t^\epsilon(x) 
\right) .
\end{equation*}
% and we assume that $\rho=\zeta*\check{\zeta}$ with $\zeta$ a rapidly decreasing function. 
% The example we have in mind is $\zeta$ (and therefore $\rho$) being
We shall prove that under the 
assumptions of Proposition~\ref{nonlocalPME to PME},
as $\epsilon\to 0$, we have
convergence to the solution to the porous medium equation with
logistic growth,
equation~\eqref{PME1}:
\begin{equation*}
	\partial_t \psi_t(x)=
	\Delta\left(\psi_t^2\right)(x)
	+\psi_t(x)\left(1-\psi_t(x)\right).
\end{equation*}

We work on the time interval $[0,T]$.
We will require a lower bound on $\int \psi_t^\epsilon(x)\log \psi_t^\epsilon(x) dx$ 
which we record as a lemma.

\begin{lemma}
\label{lower bound on ulogu}
	Suppose that there exists $\lambda\in (0,1)$ and $C<\infty$, both independent of
$\epsilon$, such that $\int \exp(\lambda |x|) \psi_0^\epsilon(x) dx<C$.
Then there exists
a constant $K<\infty$, independent of $\epsilon$, such that  
 $\int \psi_t^\epsilon(x)\log \psi_t^\epsilon (x)dx>-K$ for all $t\in [0,T]$. 
\end{lemma}
\begin{proof}
First observe that, since $x\log x$ is bounded below, 
$\int_{-1}^1\psi_t^\epsilon(x)\log\psi_t^\epsilon(x) dx$ is bounded below,
and recall that $\psi_t^\epsilon(x) \ge 0$.

@@

\begin{proof}[Proof of Proposition~\ref{nonlocalPME to PME}]
Write $\zeta = p_{\epsilon^2_\gamma / 2}$, so that $\rho = \zeta * \zeta$.
First observe that
\begin{multline*}
	\int \psi_t^\epsilon(x)\, \rho^{\epsilon}*\psi_t^\epsilon(x) dx 
=
\int\int\int \psi_t^\epsilon(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
{\zeta}^{\epsilon}(z) dz dy dx 
\\
=
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{z})
\psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z}) d\tilde{z} d\tilde{y} d \tilde{x}
	=\int\left(\zeta^{\epsilon}*\psi_t^\epsilon(x)\right)^2 dx,
\end{multline*}
where we have set $\tilde{x}=x-z$, $\tilde{y}=y-z$, $\tilde{z}=-z$.

@@

In this section we prove Theorem~\ref{thm:local_convergence},
which simultaneously scales the parameters $\theta$ and $N$ and the interaction 
kernel $\rho_F$ in the special case in which $r\equiv 1\equiv\gamma$, $q_{\theta}(x,dy)$
is isotropic with zero mean, the kernel
$\rho_F$ is Gaussian, and the scaling 
limit is a reaction-diffusion equation.
